---
title: "Bellabeat Case Study"
author: "Jenny"
date: "8/19/2021"
output: html_document
---

## Introduction and background




## Installing and loading common packages and libraries

```{r required packages}
install.packages('tidyverse', repos = "http://cran.us.r-project.org")
install.packages('ggplot2', repos = "http://cran.us.r-project.org")
```

```{r load library}
library(tidyverse)  #helps wrangle data
library(ggplot2)  #helps visualize data
getwd() #displays my working directory
setwd("//Users/jn/Case-study/CSV/Bellabeat-Case-Study") 
```


## STEP 1: COLLECT DATA

```{r}
daily_activity <- read.csv("dailyActivity_merged.csv")
sleep_day <- read.csv("sleepDay_merged.csv")
```


## STEP 2: Exploring a few key tables

Take a look at the daily_activity data.
```{r}
head(daily_activity)
```

Identify all the columsn in the daily_activity data.
```{r}
colnames(daily_activity)
```

Take a look at the sleep_day data.
```{r}
head(sleep_day)
```

Identify all the columsn in the daily_activity data.
```{r}
colnames(sleep_day)
```

Note that both datasets have the 'Id' field - this can be used to merge the datasets.

## STEP 3: Understanding some summary statistics
How many unique participants are there in each dataframe? It looks like there may be more participants in the daily activity dataset than the sleep dataset.

```{r distinct users}
n_distinct(daily_activity$Id)
n_distinct(sleep_day$Id)
```

How many observations are there in each dataframe?
```{r observations}
nrow(daily_activity)
nrow(sleep_day)
```

What are some quick summary statistics we'd want to know about each data frame?

For the daily activity dataframe:
```{r}
daily_activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes) %>%
  summary()
```

For the sleep dataframe:
```{r}
sleep_day %>%  
  select(TotalSleepRecords,
  TotalMinutesAsleep,
  TotalTimeInBed) %>%
  summary()
```

What does this tell us about how this sample of people's activities? 

## STEP 4: Plotting a few explorations

What's the relationship between steps taken in a day and sedentary minutes? How could this help inform the customer segments that we can market to? E.g. position this more as a way to get started in walking more? Or to measure steps that you're already taking?

```{r}
ggplot(data=daily_activity, aes(x=TotalSteps, y=SedentaryMinutes)) + geom_point()
```

What's the relationship between minutes asleep and time in bed? You might expect it to be almost completely linear - are there any unexpected trends?

```{r}
ggplot(data=sleep_day, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + geom_point()
```

What could these trends tell you about how to help market this product? Or areas where you might want to explore further?

## STEP 5: Merging these two datasets together

```{r}
combined_data <- merge(sleep_day, daily_activity, by="Id")
```

Take a look at how many participants are in this data set.

```{r}
n_distinct(combined_data$Id)
```

